---
title: 'Lab 1: Intro to R'
author: "Noah Collin"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  openintro::lab_report: default
---

```{r load-packages, message=T}
library(tidyverse)
library(openintro)
#library()
```

#data('arbuthnot', package='openintro')

#glimpse(arbuthnot)

### Exercise 1

```{r view-girls-counts}
arbuthnot$girls
```


### Exercise 2

It's my opinion there's not a clear trend or even direction for the population of 
girls born in this time period.  A broad summary would indicate that the number of girls born per year increased.

```{r trend-girls}
ggplot(data = arbuthnot, aes (x = year, y = girls)) +
  geom_smooth() + 
  geom_point() +
  geom_line()
```


### Exercise 3

Surprisingly, the ratio of boys baptized exceeds 50% for every data point observed.

```{r plot-prop-boys-arbuthnot}
arbuthnot <- arbuthnot %>%
  mutate(total = boys + girls)
arbuthnot <- arbuthnot %>%
  mutate(boy_ratio = boys / total)

ggplot(data = arbuthnot, aes (x = year, y = boy_ratio)) +
  geom_smooth() + 
  geom_point() +
  geom_line()
```


### Exercise 4

The "present" dataset contains data from 1940 - 2002.  That dataset has 63 rows and 3
columns.  The column names are "year", "boys", and "girls".

```{r dim-present}
present %>%
  summarize(min = min(year), max = max(year))

glimpse(present)
```


### Exercise 5

The maximum births observed in the "arbuthnot" dataset is 16145, while the minimum
total births observed in the "present" dataset is 2360399.  The scale of the numbers
is completley different.

```{r count-compare}
arbuthnot %>%
  summarize(min = min(total), max = max(total))

present <- present %>%
  mutate(total = boys + girls)

present %>%
  summarize(min = min(total), max = max(total))
```


### Exercise 6

There is indeed a greater ratio of boys born in the "present" dataset, 
just as was observed in the "arbuthnot" dataset.

```{r plot-prop-boys-present}

present <- present %>%
  mutate(boy_ratio = boys / total)

ggplot(data = present, aes (x = year, y = boy_ratio)) +
  geom_smooth() + 
  geom_point() +
  geom_line()
```


### Exercise 7

The highest number of births, 4268326 of them, occurred in 1961.

```{r find-max-total}
totalss <-  present$total

present %>%
  arrange(desc(total))

```

